---
title: "Assignment 6, Part 3"
author: "Casey Dulaney"
date: "10/25/2021"
output: html_document
---

## Load Libraries
```{r}
library(ggmap)
library(geojsonsf)
library(tmap)
library(tidyverse)
library(sf)
```

## Read Data
```{r}
pop <- read_csv("C:/Users/Casey/Documents/GEOG215.FA21/Final_project/Data/tabular/CountyPop.csv")
```

## Getting a Map

```{r}
## Read data 
nc_cty <- geojson_sf("https://opendata.arcgis.com/datasets/d192da4d0ac249fa9584109b1d626286_0.geojson")

## Print structure info
str(nc_cty)

## Map
tm_shape(nc_cty) +
  tm_polygons()


## Write as a shapefile
st_write(nc_cty, "NC_county_polygons_nconemap2.shp") 
```


## Merging
``` {r}
tab <- merge(nc_cty, pop, by.x = "CountyName", by.y = "County", all.x = "TRUE")
```

## Mapping

```{r}
tm_shape(tab) + 
  tm_polygons("ShapeSTArea", 
              style = "jenks", 
              palette = "YlOrRd",
              lwd = 0.25,
              border.col = "black",
              border.alpha = 0.5,
              title = "NC County Areas")
```



## Explanations
To spatialize this data, I used the geojsn county files along with the tabular area by county files I found online. I used the merge() function to merge the two tables, and the tmap library to make a chloropleth map. I see that the eastern part of the state has the counties with the highest areas. The west has the smallest areas. 